Geographical Python Teaching Resources: GeoPyTeR

### 2019-10-19 Geopyter.tex
Minor edits to include some new refs. Otherwise good to go as far as I'm concerned.

Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an
evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question
that teaching students and researchers how to code and how to \textit{think}
computationally presents new challenges
\citep{Etherington2016,Muller2014,rey_09jgs}. Specifically, because of the way
that these changes have unfurled, the majority of resources used in the
classroom are being developed from scratch at each institution, and often by
newly-appointed lecturers and assistant professors coming from the relatively
small number of elite research facilities active in this domain (see discussion
in \citealp{esrc2013}). Not only has this led to the duplication of effort at
multiple sites, but it has also had an enormous impact on the productivity---as
commonly measured by institutions---of early career researchers, many of whom
are in their first teaching post.

@@

With these criteria and resources in mind, we began looking for ways to lower
the `entry costs' for faculty to both take from, and share into, the corpus of
instructional material, as well as for a way to address the
customisation/localisation challenge. \gp was our solution: a single, open hub
of teaching concepts and practical programming tasks that can be flexibly
assembled into classes, or entire courses! In the \gp approach we recognise that
teachers need to be able to develop their own classroom `story' at their own
pace (and in conformance with their institution's teaching pattern), while
supporting them with a library of up-to-date materials from which they can pick
and choose. So in much the same way that a computer application is compiled from
the contributions of many developers, \gp allows individual classes and entire
courses to be `compiled' from the contributions of many skilled teachers and
developers.

@@

At least in our experience, the pace of technological and methodological change
is such that the gap between what instructors know and what students know is
substantial, and in many cases might even be growing. The differences within
individual student cohorts may be greater still, however, and while students
need to be able to begin their learning `journey' at some arbitrary point in the
progression from `learning to code' towards `computational social scientist'
\citep{Lazer2009}, instructors require the ability to assemble a course that
meets students where they \emph{are} rather than where the instructor or
institution might \emph{wish} them to be.

>>>>

Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an
evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question
that teaching students and researchers how to code and how to \textit{think}
computationally presents new challenges
\citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}). Consequently, the majority of resources used in the
classroom are being developed from scratch at each institution, and often by
newly-appointed lecturers and assistant professors coming from the relatively
small number of elite research facilities active in this domain (see discussion
in \citealp{esrc2013}). Not only has this led to the duplication of effort at
multiple sites, but it has also had an enormous impact on the productivity---as
commonly measured by institutions---of early career researchers, many of whom
are in their first teaching post.

@@

With these criteria and resources in mind, we began looking for ways to lower
the `entry costs' for faculty to both take from, and share into, the corpus of
instructional material, as well as for a way to address the
customisation/localisation challenge. \gp was our solution: a single, open hub
of teaching concepts and practical programming tasks that can be flexibly
assembled into classes, or entire courses! In the \gp approach we recognise that
teachers need to be able to develop their own classroom `story' at their own
pace (and in conformance with their institution's teaching pattern), while
supporting them with a library of up-to-date materials from which they can pick
and choose. So in much the same way that a computer application is compiled from
the contributions of many developers, \gp allows individual classes and entire
courses to be `compiled' from the contributions of many skilled teachers and
developers. As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[][\eg]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!

@@

At least in our experience, the pace of technological and methodological change
is such that the gap between what instructors know and what students know is
substantial, and in many cases might even be growing. The differences within
individual student cohorts may be greater still, however, and while students
need to be able to begin their learning `journey' at some arbitrary point in the
progression from `learning to code' towards `computational social scientist'
\citep{Lazer2009}, instructors require the ability to assemble a course that
meets students where they \emph{are} rather than where the instructor or
institution might \emph{wish} them to be. At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid.
